---
title: "biRdbRained"
author: "Elizabeth Breitmeyer, Ana Pirosca, Georgia Sisk, Letacianna Stoudmire"
date: "2023-11-29"
output: html_document
bibliography: BIOL3140.bib
---

## Introduction

Climate change is a phenomenon that is dramatically impacting the natural world. Anthropogenic climate change has led to increases in mean annual global temperature which can have catastrophic impacts on the biosphere. For these reasons, it is important to investigate how climate change and shifting global temperatures are affecting different species.

Birds, as a group, have been extensively studied by scientists and civilians alike. In particular, migrating birds have garnered much attention. In the spring in the northern hemisphere, many neotropical passerines take a long-distance migration from Central/South America to higher, more temperate latitudes in order to forage and breed. Some of these species fly over the Gulf of Mexico and are referred to as trans-Gulf migrants (TGMs).The date that these migrants arrive at these breeding grounds is extremely important in terms of fitness. With climate change leading to warmer springs, it may be important for TGMs to shift their arrival date at breeding grounds or they may face population declines.

Previous research by @miller2008bird used 33 years of bird capture data from Manomet, Massachusetts to assess variation in spring migration times for 32 species of North American passerines. They found that migration times of many species occurred earlier over time, and several species' migration times were correlated with spring temperatures. Similarly, we seek to study the effect of local weather conditions on the arrival time of five species of trans-Gulf migrants in Massachusetts. The following five species were studied:

-   Belted kingfisher, *Megaceryle alcyon*
-   Common nighthawk, *Chordeiles minor*
-   Marsh wren, *Cistothorus palustris*
-   Olive-sided flycatcher, *Contopus cooperi*
-   Scarlet tanager, *Piranga olivacea*

Ultimately our study sought to answer the following questions:

1.  Does arrival time vary according to temperature and wind variables along migration route for trans-Gulf migrants migrating to Massachusetts?
2.  If arrival time does vary with meteorological conditions, what role will climate change potentially play in the population status of TGMs arriving in MA during the spring?
3.  What role do climatic variables play in long-distance migration in passerine birds?

## Methods

Occurrence data for five selected species of TGMs was gathered using the rgbif package. Logistic modeling operations were then used to predict the Julian day on which 25% of the population of the species arrives each year. The NOAA's NCDC API enabled identification of weather stations along the TGMs' migration path to Boston, and weather data from each station was collected. The effect of weather parameters on the arrival time was evaluated using linear mixed-effect modeling from the lme4 package. The car package was then used to perform an anova analysis of the interactions in order to determine the significance of the fixed effects.

## Results

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rgbif)
library(tidyverse)
library(MuMIn)
library(rnoaa)
library(data.table)
library(ggmap)
library(usmap)
library(magick)
library(cowplot)
library(lme4)
library(car)
library(data.table)
library(knitr)
library(kableExtra)
library(sp)
library(sf)
```

```{r, echo=FALSE}
kingfisher <- occ_data(scientificName = "Megaceryle alcyon", stateProvince="Massachusetts", limit=200,year=2019)
nighthawk <- occ_data(scientificName = "Chordeiles minor", stateProvince="Massachusetts", limit=200,year=2019)
wren <- occ_data(scientificName = "Cistothorus palustris", stateProvince="Massachusetts", limit=200,year=2019)
flycatcher <- occ_data(scientificName = "Contopus cooperi", stateProvince="Massachusetts", limit=200,year=2019)
tanager <- occ_data(scientificName = "Piranga olivacea", stateProvince="Massachusetts", limit=200,year=2019)

MA<- map_data('state', 'massachusetts')

kingfisher.p <- ggplot(MA, aes(long,lat,group=subregion) )+
geom_polygon(colour = "gray",fill="gray90")+geom_point(data=kingfisher[[2]],aes(x=decimalLongitude,y=decimalLatitude,size=individualCount),alpha=0.3,inherit.aes = F)+ coord_quickmap()+theme_void()
nighthawk.p <- ggplot(MA, aes(long,lat,group=subregion) )+
geom_polygon(colour = "gray",fill="gray90")+geom_point(data=nighthawk[[2]],aes(x=decimalLongitude,y=decimalLatitude,size=individualCount),alpha=0.3,inherit.aes = F)+ coord_quickmap()+theme_void()
wren.p <- ggplot(MA, aes(long,lat,group=subregion) )+
geom_polygon(colour = "gray",fill="gray90")+geom_point(data=wren[[2]],aes(x=decimalLongitude,y=decimalLatitude,size=individualCount),alpha=0.3,inherit.aes = F)+ coord_quickmap()+theme_void()
flycatcher.p <- ggplot(MA, aes(long,lat,group=subregion) )+
geom_polygon(colour = "gray",fill="gray90")+geom_point(data=flycatcher[[2]],aes(x=decimalLongitude,y=decimalLatitude,size=individualCount),alpha=0.3,inherit.aes = F)+ coord_quickmap()+theme_void()
tanager.p <- ggplot(MA, aes(long,lat,group=subregion) )+
geom_polygon(colour = "gray",fill="gray90")+geom_point(data=tanager[[2]],aes(x=decimalLongitude,y=decimalLatitude,size=individualCount),alpha=0.3,inherit.aes = F)+ coord_quickmap()+theme_void()

#kingfisher
kingfisher.p2 <- ggdraw()+ draw_plot(kingfisher.p)
print(kingfisher.p2)

#nighthawk
nighthawk.p2 <- ggdraw()+ draw_plot(nighthawk.p)
print(nighthawk.p2)

#wren
wren.p2 <- ggdraw()+ draw_plot(wren.p)
print(wren.p2)

#flycatcher
flycatcher.p2 <- ggdraw()+ draw_plot(flycatcher.p)
print(flycatcher.p2)

#tanager
tanager.p2 <- ggdraw()+ draw_plot(tanager.p)
print(tanager.p2)
```

```{r, eval=FALSE,echo=FALSE,warning=FALSE, message=FALSE}
species <- c("Megaceryle alcyon","Chordeiles minor","Cistothorus palustris","Contopus cooperi","Piranga olivacea")
y <- paste0("1999",",","2019")
m <- paste0("3",",","6")
dat.l <-list()
for(s in species){
n.obs <-occ_data(scientificName = s, year = y, month = m, limit = 0, country = "US", basisOfRecord = "HUMAN_OBSERVATION", stateProvince="Massachusetts")$meta$count
print(n.obs)
dat.l[[paste0(s)]] <- occ_data(scientificName = s, year = y, month = m, 
limit = n.obs, country = "US",
basisOfRecord = "HUMAN_OBSERVATION",
stateProvince = "Massachusetts")[[2]]
}
```

```{r, eval=FALSE,include=FALSE,warning=FALSE,message=FALSE}
saveRDS(dat,"massbird.data.RDS")
```

```{r species count vs year,echo=FALSE,warning=FALSE,message=FALSE, fig.cap= "**Fig. 2.** Occurence of five TGM species in Massachusetts from 1990 to 2019."}
dat <- bind_rows(dat.l, .id = "species")
dat <- readRDS("massbird.data.RDS")%>%
  group_by(year, species)%>%
  summarise(count=sum(individualCount,na.rm = T))%>%
  ggplot(aes(x=year,y=count,col=species))+geom_point()
```

```{r, include=FALSE}
options(noaakey = "AlfFQdBUsXZVspOWUYsgIjhNEttAQaYa")
sts <- c(
"GHCND:USW00013894", #Mobible, AL 2k away about 10 days away @200 km/day
"GHCND:USW00013881", #Charlotte, NC 1000 km away about 6 days away @200 km/day
"GHCND:USW00014739" #Boston
)
bos <- ncdc_stations(stationid = "GHCND:USW00014739")
print(bos)
sta.d <- bind_rows( #bind the rows
  lapply(sts,function(x) ncdc_stations(stationid = x)$data ) #use lapply to run through stations
  )%>%
  mutate(usmap_transform(.,input_names = c("longitude","latitude"),output_names = c("longitude.1", "latitude.1")))%>% 
  mutate(name=str_sub(name, -5,-4))%>%
  mutate(migr.day=c(10,5,0))%>% 
  separate(id,into = c("station.type","id"))%>%
        print()
```

```{r}
plot_usmap(
include = c(.northeast_region,.south_region,.east_north_central)
)+geom_point(data=sta.d,aes(x=longitude.1,y=latitude.1,col=name),size=5)+geom_label(data=sta.d,aes(x=longitude.1,y=latitude.1,col=name,label=name),size=5,nudge_x = 1e6*0.25)+theme(legend.position = "none")
```

```{r, include=FALSE,cache=TRUE}
weather.d <- meteo_pull_monitors(sta.d$id,date_min = "2000-01-01")
head(weather.d)
```

### Belted kingfisher (Megaceryle alcyon)

```{r, echo=FALSE,warning=FALSE,message=FALSE}
mc<- dat%>%
  filter(species=="Megaceryle alcyon")%>%
  group_by(year)%>%
  mutate(
    date=as.Date(paste0(year,"-",month,"-",day)),
    j.day=julian(date,origin=as.Date(paste0(unique(year),"-01-01"))))%>%
  group_by(species,year,j.day,date)%>%
  summarise(day.tot=sum(individualCount,na.rm=T))%>%
  group_by(species,year)%>%
  mutate(prop=cumsum(day.tot/sum(day.tot,na.rm = T)))%>%
  filter(year>1999)

mc%>% 
  ggplot(aes(j.day,prop))+geom_point()+facet_wrap(year~.)
mc.pred <- mc%>%
  group_by(year)%>%
  summarize(
    pred=predict(nls(prop~SSlogis(j.day,Asym, xmid, scal)),newdata=data.frame(j.day=min(j.day):max(j.day))),
j.day=min(j.day):max(j.day),
)%>%
left_join(mc%>%dplyr::select(j.day,date))
```

```{r, fig.cap= "Proportion of the population of Belted kingfisher that has arrived in Boston, MA by the given Julian day. Logistic models plotted in blue to estimate the arrival date."}
mc%>%
ggplot(aes(j.day,prop))+geom_point(aes=0.3)+geom_line(data=mc.pred,aes(x=j.day,y=pred),col="blue",size=2)+facet_wrap(year~.)
```

```{r, fig.cap= "Julian day from each year corresponding to the arrival of 25% of the Belted kingfisher population to Boston, MA."}
mc.arrive.date <-mc.pred%>%
  group_by(year)%>%
  filter(j.day==j.day[which.min(abs(pred-0.25))])
mc.arrive.date%>%
  ggplot(aes(year,j.day))+geom_point()
```

```{r, include=FALSE}
weather.d <- weather.d%>%
  mutate(year=as.integer(str_sub(date,1,4)), #add year
date=as.Date(date))%>%
group_by(year)%>% #group by year so we can compute julian day
 mutate(j.day=julian(date,origin=as.Date(paste0(unique(year),"-01-01"))), #add julian day
date2=date,
wdir.rad=(180-abs(wdf2-180))*pi/180, #radians so we can use a trig function to compute wind vector, scale degrees first to 180 scale to 2x pi and subtract from 180 (wind comes out of a direction)
wvec=cos(wdir.rad)*-1*awnd # we want a negative value for positive value for 2x pi
)%>% #store day in new column
dplyr::select(id,year,date2,j.day,tmin,tmax,wvec)%>% #select the rows we need
left_join(sta.d%>%select(id,name,migr.day))%>% #add the station id info (ie. name)
mutate(j.day=j.day+migr.day)#make j.day ahead of BOS according to the migration days away so we can join weather along path
mc.arr.weath <- mc.arrive.date%>%
left_join(weather.d)%>%
left_join(mc%>%dplyr::select(year,date,j.day))
head(mc.arr.weath)
weather.wk <-weather.d %>% 
group_by(year,name) %>% 
mutate(wk.tmin = frollmean(tmin, n=14,align="right"),
wk.tmax = frollmean(tmax, n=14,align="right"),
wk.wvec = frollmean(wvec, n=14,align="right")
)%>%
dplyr::select(j.day,date2,name,wk.tmin,wk.tmax,wk.wvec)
mc.arr.weath2 <- mc.arrive.date%>%
left_join(weather.wk)
head(mc.arr.weath2)
mc.lmer <- lmer(j.day~tmin*tmax*wvec+(1|name),mc.arr.weath,na.action = "na.fail")
Anova(mc.lmer)
mc.lmer2 <- lmer(j.day~wk.tmin*wk.tmax*wk.wvec+(1|name),mc.arr.weath2,na.action = "na.fail")
Anova(mc.lmer2) 
mc.arr.aic <- dredge(mc.lmer2,fixed = c("wk.tmin","wk.tmax","wk.wvec"),)
mc.kb <- kable(mc.arr.aic[1:4,],caption = "Fit values for nested models of the most complicated linear mixed-effect model.")
kable_styling(mc.kb)
best.lmer <-lmer(j.day~wk.tmin+wk.tmax+wk.wvec+(1|name),mc.arr.weath2,na.action = "na.fail")
Anova(best.lmer)
```

### Common nighthawk (Chordeiles minor)

```{r, echo=FALSE,warning=FALSE,message=FALSE}
mc<- dat%>% 
  filter(species=="Chordeiles minor")%>%
  group_by(year)%>%
  mutate(date=as.Date(paste0(year,"-",month,"-",day)),
         j.day=julian(date,origin=as.Date(paste0(unique(year),"-01-01")))
         )%>%
  group_by(species,year,j.day,date)%>%
  summarise(day.tot=sum(individualCount,na.rm=T))%>%
  group_by(species,year)%>%
  mutate(prop=cumsum(day.tot/sum(day.tot,na.rm = T)))%>%
  filter(year>1999)
mc%>%
  ggplot(aes(j.day,prop))+geom_point()+facet_wrap(year~.)
mc.pred <- mc%>%
  group_by(year)%>%
  summarize(pred=predict(nls(prop~SSlogis(j.day,Asym, xmid, scal)),newdata=data.frame(j.day=min(j.day):max(j.day))),
j.day=min(j.day):max(j.day),
)%>%
left_join(mc%>%dplyr::select(j.day,date))
```

```{r, fig.cap= "Proportion of the population of Common nighthawk that has arrived in Boston, MA by the given Julian day. Logistic models plotted in blue to estimate the arrival date."}
mc%>%
ggplot(aes(j.day,prop))+geom_point(aes=0.3)+geom_line(data=mc.pred,aes(x=j.day,y=pred),col="blue",size=2)+facet_wrap(year~.)
```

```{r, fig.cap= "Julian day from each year corresponding to the arrival of 25% of the Common nighthawk population to Boston, MA."}
mc.arrive.date <-mc.pred%>%
  group_by(year)%>%
  filter(j.day==j.day[which.min(abs(pred-0.25))])
mc.arrive.date%>%
  ggplot(aes(year,j.day))+geom_point()
```

```{r, include=FALSE}
mc.arr.weath <- mc.arrive.date%>%
  left_join(weather.d)%>%
  left_join(mc%>%dplyr::select(year,date,j.day))
head(mc.arr.weath)
weather.wk <-weather.d %>% 
  group_by(year,name) %>% 
  mutate(wk.tmin = frollmean(tmin, n=14,align="right"),
         wk.tmax = frollmean(tmax, n=14,align="right"),
         wk.wvec = frollmean(wvec, n=14,align="right")
)%>%
dplyr::select(j.day,date2,name,wk.tmin,wk.tmax,wk.wvec)
mc.arr.weath2 <- mc.arrive.date%>%
  left_join(weather.wk)
head(mc.arr.weath2)
mc.lmer <- lmer(j.day~tmin*tmax*wvec+(1|name),mc.arr.weath,na.action = "na.fail")
Anova(mc.lmer)
mc.lmer2 <- lmer(j.day~wk.tmin*wk.tmax*wk.wvec+(1|name),mc.arr.weath2,na.action = "na.fail")
Anova(mc.lmer2) 
mc.arr.aic <- dredge(mc.lmer2,fixed = c("wk.tmin","wk.tmax","wk.wvec"),)
mc.kb <- kable(mc.arr.aic[1:4,],caption = "Fit values for nested models of the most complicated linear mixed-effect model.")
kable_styling(mc.kb)
best.lmer <-lmer(j.day~wk.tmin+wk.tmax+wk.wvec+(1|name),mc.arr.weath2,na.action = "na.fail")
Anova(best.lmer)
```

### Marsh wren 
```{r, echo=FALSE,warning=FALSE,message=FALSE}
mc<- dat%>% 
  filter(species=="Cistothorus palustris")%>%
  group_by(year)%>%
  mutate(date=as.Date(paste0(year,"-",month,"-",day)),
         j.day=julian(date,origin=as.Date(paste0(unique(year),"-01-01")))
)%>%
group_by(species,year,j.day,date)%>%
  summarise(day.tot=sum(individualCount,na.rm=T))%>%
  group_by(species,year)%>%
  mutate(prop=cumsum(day.tot/sum(day.tot,na.rm = T)))%>%
  filter(year>1999)
mc%>% 
  ggplot(aes(j.day,prop))+geom_point()+facet_wrap(year~.)
mc.pred <- mc%>%
  group_by(year)%>%
  summarize(
    pred=predict(nls(prop~SSlogis(j.day,Asym, xmid, scal)),newdata=data.frame(j.day=min(j.day):max(j.day))),
j.day=min(j.day):max(j.day),
)%>%
left_join(mc%>%dplyr::select(j.day,date))
```

```{r, fig.cap= "Proportion of the population of Belted kingfisher that has arrived in Boston, MA by the given Julian day. Logistic models plotted in blue to estimate the arrival date."}
mc%>%
ggplot(aes(j.day,prop))+geom_point(aes=0.3)+geom_line(data=mc.pred,aes(x=j.day,y=pred),col="blue",size=2)+facet_wrap(year~.)
```

```{r, fig.cap= "Julian day from each year corresponding to the arrival of 25% of the Belted kingfisher population to Boston, MA."}
mc.arrive.date <-mc.pred%>%
  group_by(year)%>%
  filter(j.day==j.day[which.min(abs(pred-0.25))])
mc.arrive.date%>%
  ggplot(aes(year,j.day))+geom_point()
```

```{r, include=FALSE}
weather.d <- weather.d%>%
  mutate(year=as.integer(str_sub(date,1,4)), #add year
date=as.Date(date))%>%
group_by(year)%>% #group by year so we can compute julian day
 mutate(j.day=julian(date,origin=as.Date(paste0(unique(year),"-01-01"))), #add julian day
date2=date,
wdir.rad=(180-abs(wdf2-180))*pi/180, #radians so we can use a trig function to compute wind vector, scale degrees first to 180 scale to 2x pi and subtract from 180 (wind comes out of a direction)
wvec=cos(wdir.rad)*-1*awnd # we want a negative value for positive value for 2x pi
)%>% #store day in new column
dplyr::select(id,year,date2,j.day,tmin,tmax,wvec)%>% #select the rows we need
left_join(sta.d%>%select(id,name,migr.day))%>% #add the station id info (ie. name)
mutate(j.day=j.day+migr.day)#make j.day ahead of BOS according to the migration days away so we can join weather along path
mc.arr.weath <- mc.arrive.date%>%
left_join(weather.d)%>%
left_join(mc%>%dplyr::select(year,date,j.day))
head(mc.arr.weath)
weather.wk <-weather.d %>% 
group_by(year,name) %>% 
mutate(wk.tmin = frollmean(tmin, n=14,align="right"),
wk.tmax = frollmean(tmax, n=14,align="right"),
wk.wvec = frollmean(wvec, n=14,align="right")
)%>%
dplyr::select(j.day,date2,name,wk.tmin,wk.tmax,wk.wvec)
mc.arr.weath2 <- mc.arrive.date%>%
left_join(weather.wk)
head(mc.arr.weath2)
mc.lmer <- lmer(j.day~tmin*tmax*wvec+(1|name),mc.arr.weath,na.action = "na.fail")
Anova(mc.lmer)
mc.lmer2 <- lmer(j.day~wk.tmin*wk.tmax*wk.wvec+(1|name),mc.arr.weath2,na.action = "na.fail")
Anova(mc.lmer2) 
mc.arr.aic <- dredge(mc.lmer2,fixed = c("wk.tmin","wk.tmax","wk.wvec"),)
mc.kb <- kable(mc.arr.aic[1:4,],caption = "Fit values for nested models of the most complicated linear mixed-effect model.")
kable_styling(mc.kb)
best.lmer <-lmer(j.day~wk.tmin+wk.tmax+wk.wvec+(1|name),mc.arr.weath2,na.action = "na.fail")
Anova(best.lmer)
```
### Olive-sided flycatcher
```{r, echo=FALSE,warning=FALSE,message=FALSE}
mc<- dat%>% 
  filter(species=="Contopus cooperi")%>%
  group_by(year)%>%
  mutate(date=as.Date(paste0(year,"-",month,"-",day)),
         j.day=julian(date,origin=as.Date(paste0(unique(year),"-01-01")))
)%>%
group_by(species,year,j.day,date)%>%
  summarise(day.tot=sum(individualCount,na.rm=T))%>%
  group_by(species,year)%>%
  mutate(prop=cumsum(day.tot/sum(day.tot,na.rm = T)))%>%
  filter(year>1999)
mc%>% 
  ggplot(aes(j.day,prop))+geom_point()+facet_wrap(year~.)
mc.pred <- mc%>%
  group_by(year)%>%
  summarize(
    pred=predict(nls(prop~SSlogis(j.day,Asym, xmid, scal)),newdata=data.frame(j.day=min(j.day):max(j.day))),
j.day=min(j.day):max(j.day),
)%>%
left_join(mc%>%dplyr::select(j.day,date))
```

```{r, fig.cap= "Proportion of the population of Belted kingfisher that has arrived in Boston, MA by the given Julian day. Logistic models plotted in blue to estimate the arrival date."}
mc%>%
ggplot(aes(j.day,prop))+geom_point(aes=0.3)+geom_line(data=mc.pred,aes(x=j.day,y=pred),col="blue",size=2)+facet_wrap(year~.)
```

```{r, fig.cap= "Julian day from each year corresponding to the arrival of 25% of the Belted kingfisher population to Boston, MA."}
mc.arrive.date <-mc.pred%>%
  group_by(year)%>%
  filter(j.day==j.day[which.min(abs(pred-0.25))])
mc.arrive.date%>%
  ggplot(aes(year,j.day))+geom_point()
```

```{r, include=FALSE}
weather.d <- weather.d%>%
  mutate(year=as.integer(str_sub(date,1,4)), #add year
date=as.Date(date))%>%
group_by(year)%>% #group by year so we can compute julian day
 mutate(j.day=julian(date,origin=as.Date(paste0(unique(year),"-01-01"))), #add julian day
date2=date,
wdir.rad=(180-abs(wdf2-180))*pi/180, #radians so we can use a trig function to compute wind vector, scale degrees first to 180 scale to 2x pi and subtract from 180 (wind comes out of a direction)
wvec=cos(wdir.rad)*-1*awnd # we want a negative value for positive value for 2x pi
)%>% #store day in new column
dplyr::select(id,year,date2,j.day,tmin,tmax,wvec)%>% #select the rows we need
left_join(sta.d%>%select(id,name,migr.day))%>% #add the station id info (ie. name)
mutate(j.day=j.day+migr.day)#make j.day ahead of BOS according to the migration days away so we can join weather along path
mc.arr.weath <- mc.arrive.date%>%
left_join(weather.d)%>%
left_join(mc%>%dplyr::select(year,date,j.day))
head(mc.arr.weath)
weather.wk <-weather.d %>% 
group_by(year,name) %>% 
mutate(wk.tmin = frollmean(tmin, n=14,align="right"),
wk.tmax = frollmean(tmax, n=14,align="right"),
wk.wvec = frollmean(wvec, n=14,align="right")
)%>%
dplyr::select(j.day,date2,name,wk.tmin,wk.tmax,wk.wvec)
mc.arr.weath2 <- mc.arrive.date%>%
left_join(weather.wk)
head(mc.arr.weath2)
mc.lmer <- lmer(j.day~tmin*tmax*wvec+(1|name),mc.arr.weath,na.action = "na.fail")
Anova(mc.lmer)
mc.lmer2 <- lmer(j.day~wk.tmin*wk.tmax*wk.wvec+(1|name),mc.arr.weath2,na.action = "na.fail")
Anova(mc.lmer2) 
mc.arr.aic <- dredge(mc.lmer2,fixed = c("wk.tmin","wk.tmax","wk.wvec"),)
mc.kb <- kable(mc.arr.aic[1:4,],caption = "Fit values for nested models of the most complicated linear mixed-effect model.")
kable_styling(mc.kb)
best.lmer <-lmer(j.day~wk.tmin+wk.tmax+wk.wvec+(1|name),mc.arr.weath2,na.action = "na.fail")
Anova(best.lmer)
```

###Scarlet tanager 
```{r, echo=FALSE,warning=FALSE,message=FALSE}
mc<- dat%>% 
  filter(species=="Piranga olivacea")%>%
  group_by(year)%>%
  mutate(date=as.Date(paste0(year,"-",month,"-",day)),
         j.day=julian(date,origin=as.Date(paste0(unique(year),"-01-01")))
)%>%
group_by(species,year,j.day,date)%>%
  summarise(day.tot=sum(individualCount,na.rm=T))%>%
  group_by(species,year)%>%
  mutate(prop=cumsum(day.tot/sum(day.tot,na.rm = T)))%>%
  filter(year>1999)
mc%>% 
  ggplot(aes(j.day,prop))+geom_point()+facet_wrap(year~.)
mc.pred <- mc%>%
  group_by(year)%>%
  summarize(
    pred=predict(nls(prop~SSlogis(j.day,Asym, xmid, scal)),newdata=data.frame(j.day=min(j.day):max(j.day))),
j.day=min(j.day):max(j.day),
)%>%
left_join(mc%>%dplyr::select(j.day,date))
```

```{r, fig.cap= "Proportion of the population of Belted kingfisher that has arrived in Boston, MA by the given Julian day. Logistic models plotted in blue to estimate the arrival date."}
mc%>%
ggplot(aes(j.day,prop))+geom_point(aes=0.3)+geom_line(data=mc.pred,aes(x=j.day,y=pred),col="blue",size=2)+facet_wrap(year~.)
```

```{r, fig.cap= "Julian day from each year corresponding to the arrival of 25% of the Belted kingfisher population to Boston, MA."}
mc.arrive.date <-mc.pred%>%
  group_by(year)%>%
  filter(j.day==j.day[which.min(abs(pred-0.25))])
mc.arrive.date%>%
  ggplot(aes(year,j.day))+geom_point()
```

```{r, include=FALSE}
weather.d <- weather.d%>%
  mutate(year=as.integer(str_sub(date,1,4)), #add year
date=as.Date(date))%>%
group_by(year)%>% #group by year so we can compute julian day
 mutate(j.day=julian(date,origin=as.Date(paste0(unique(year),"-01-01"))), #add julian day
date2=date,
wdir.rad=(180-abs(wdf2-180))*pi/180, 
wvec=cos(wdir.rad)*-1*awnd 
)%>% #store day in new column
dplyr::select(id,year,date2,j.day,tmin,tmax,wvec)%>% 
left_join(sta.d%>%select(id,name,migr.day))%>% 
mutate(j.day=j.day+migr.day)#make j.day ahead of BOS according to the migration days away so we can join weather along path
mc.arr.weath <- mc.arrive.date%>%
left_join(weather.d)%>%
left_join(mc%>%dplyr::select(year,date,j.day))
head(mc.arr.weath)
weather.wk <-weather.d %>% 
group_by(year,name) %>% 
mutate(wk.tmin = frollmean(tmin, n=14,align="right"),
wk.tmax = frollmean(tmax, n=14,align="right"),
wk.wvec = frollmean(wvec, n=14,align="right")
)%>%
dplyr::select(j.day,date2,name,wk.tmin,wk.tmax,wk.wvec)
mc.arr.weath2 <- mc.arrive.date%>%
left_join(weather.wk)
head(mc.arr.weath2)
mc.lmer <- lmer(j.day~tmin*tmax*wvec+(1|name),mc.arr.weath,na.action = "na.fail")
Anova(mc.lmer)
mc.lmer2 <- lmer(j.day~wk.tmin*wk.tmax*wk.wvec+(1|name),mc.arr.weath2,na.action = "na.fail")
Anova(mc.lmer2) 
mc.arr.aic <- dredge(mc.lmer2,fixed = c("wk.tmin","wk.tmax","wk.wvec"),)
mc.kb <- kable(mc.arr.aic[1:4,],caption = "Fit values for nested models of the most complicated linear mixed-effect model.")
kable_styling(mc.kb)
best.lmer <-lmer(j.day~wk.tmin+wk.tmax+wk.wvec+(1|name),mc.arr.weath2,na.action = "na.fail")
Anova(best.lmer)
```

## Discussion
Our results demonstrate that TGM arrival time in MA exhibits a significant relationship with temperature and wind variables along the migration route. Specifically, earlier arrivals were associated with warmer temperatures and tailwinds. This finding aligns with previous studies that have documented the impact of temperature and wind on avian migration timing.

This finding aligns with previous studies that have documented the impact of temperature and wind on avian migration timing. For instance, @marra2005 observed a positive correlation between spring arrival dates of warblers in North America and pre-migratory temperatures in their wintering grounds, suggesting that warmer temperatures facilitate earlier departures and potentially faster journeys. Similarly, @Hedenstrom1992 reported that tailwinds during migratory flights significantly reduced travel time for several shorebird species, highlighting the wind's role in facilitating efficient movement.

Our study underscores the urgency of understanding how climate change will impact long-distance migratory birds like TGMs. By integrating meteorological data with avian migration data, we can develop robust predictive models that inform conservation efforts and help mitigate the potential negative consequences of a changing climate on these vulnerable populations.

## Author Contributions

Elizabeth: Introduction section

Ana: Methods and Discussion

Georgia: Methods section and Results section

Letacianna: Results section

## References
